Parameteric controls on responses to biogeochemical forcing in the CLM5

### 2018-11-19 main.tex

\begin{document}

\authors{Rosie Fisher\affil{1},
William Wieder\affil{1},
Benjamin Sanderson\affil{2},
Charles Koven\affil{3},
Keith Oleson\affil{1},
Chonggang Xu\affil{4},
Ashehad Ali\affil{1},
Joshua Fisher\affil{5},
Mingjie Shi\affil{5},
Anthony Walker\affil{6},
Gordon Bonan\affil{1},
David M. Lawrence\affil{1}
}

@@

For this experiment we utilized six different experimental sites. Two in tropical forests (Caxiuana and Barro Colorado Island) two in temperate forest (the Oak Ridge National Lab - Free Air Carbon Dioxide Enrichment (FACE) experiment and Harvard Forest), one in high altitude forest (Niwot Ridge, Colorado) and one in seasonally dry evergreen forest (Metolius Forest) **tbd**.

@@

\subsection{Perturbed Physics Ensemble (PPE)}
The CLM5 has a large array of parameters that might have direct and indirect impacts on biogeochemical cycling in the CLM5. For this manuscript, we focus on parameters included in recent developments in the vegetation model, and how they might impact upon the responses to CO$_{2}$ and N fertilization. Therefore, 
we focus on parameters of particular relevance to the C and N cycles and on those which are newly introduced by the modified N cycle. The choice of parameters used here is informed primarily by the model structure, as well as many iterative and informal investigations of the CLM5 model during the development process.

@@

\subsubsection{Parameter Choices}

@@

\subsubsection{Parameters not used}

We conducted a set of tests on the parameters of the denitrification model, to assess the impacts of differing rates of N loss on the system. However, recent analysis has shown that the rates of overall denitrification in CLM5 are very low, overall (compared to plant uptake and immobilization \textbf{I've forgotten why this is? Can someone remind me}) and hence these parameters had very little impact. We did not conduct tests on leaf lifespan, since for evergreen trees in CLM5, leaf lifespan and specific leaf area have very similar impacts on overall achievable leaf area, and for deciduous trees, leaf lifespan is primarily determined by phenological triggers. We did not investigate any parameters relating to the crop model (Lombardozzi et al. in prep) nor to plant harvest, fire, dust, or soil decomposition, (the latter is largely unaltered from CLM4.5 as per \cite{koven2013}.

@@

\emph{Specific Leaf Area} and \emph{leafcn} are defined in the TRY plant trait database(\url{www.try-db.org} and we take their values from the analysis by (\cite{kattge2011}) of their mean and distribution. For both parameters, values are log-normally distributed. We here take the PFT average log-normalised standard deviation range from TRY and apply it across all PFTs (thus the lower bounds are closer to the mean than the higher bounds, Table \ref{table_ranges}). The range of \emph{fracfixers} and \emph{frac\_ecto} were set to vary across the whole logical range from 0 to 1 (0 to 100\% of ecosystem productivity being available for fixation or ectomycorrhizal uptake). \emph{grperc} ranges from its logical lower bound of zero up to the range of the previously used CLM value (0.3). \emph{lmr\_intercept} and \emph{medlyn\_slope} are both set such that the range of values tested represents the range across all the PFTs reported in \cite{atkin2015} and \cite{dekauwe2015}. The ranges of the \emph{cn\_flex} parameters are unknown, since these physiological controls are poorly understood. Thus, we explore an order of magnitude in each direction from the default, to obtain an understanding of the general sensitivity to this parameterization (thus, these results should not be interpreted as covering a constrained range of variation). The same issues apply to $N_{costs}$.

@@

\subsubsection{Low leaf area investigations}
In general, we found that leaf area indices predicted by the CLM5 default parameter set were often low  compared to published estimates for these sites. (for Caxiuana, 3.0 v.s. 5.3 \cite{fisher2007}, for Niwot Ridge, 1.7 vs. 4.2 \cite{bowling2009}, for Harvard Forest 2.6 vs. 3.5 \cite{williams1996} and for Metolius, 0.5 (modeled) v.s. 1.5 (\cite{spadavecchia2011}). This is a somewhat surprising result, given the generally good agreement of globally-driven model output with LAI reported by \cite{lawrence2018}, and may plausibly reflect biases between the use of reanalysis products used to drive the model and site-specific meteorological data.

To interrogate this further, we ran a separate set of simulations with higher allocation to leaves (generated from lower stem allocation parameters). These indicated that, with the resulting higher LAI's, the model was less sensitive to both CO$_{2}$ and N deposition, given the amplifying effect of increasing leaf area increasing fluxes at unsaturated LAIs.   For example, at Harvard Forest, the high leaf allocation  simulation (LAI of 4.6, vs 2.6 for the default parameters) had a lower CO$_{2}$ response (1.43, 1.66, 1.47 for GPP, NPP and LAI in the control vs. 1.18, 1.17 and 1.22 for the high LAI case) and a lower N response as well (1.1, 1.26, 1.21 for the control GPP, NPP, LAI, and 1.05, 1.20 and 1.15 for the high LAI case). Similar results were found at all other sites.  Thus, by virtue of their lower LAI, the overall fertilization rates at these sites may be substantially larger than those in the model driven with reanalysis meteorology (analysis tbd).

\subsubsection{Water and N limitation across sites}
Figure \ref{btran state} illustrates the relative water and nitrogen limitation in the control (unfertilized) parameter ensemble. Water limitation in CLM5 can be expressed via the $B_{tran,mn}$ parameter, which is a model property calculated as a function of prognostic leaf water potential, which impacts gas exchange processes in a similar manner to the $B_{tran}$ parameter in CLM4.5, (in that it is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$). Because leaf water potential changes diurnally, we now use the minimum term to calculate relative water stress between sites. For N limitation, the fraction of total NPP which is used to pay for nitrogen uptake, is a direct indicator of how N stress affects the carbon budget.

The Caxiuana site experienced substantial drought stress in these simulations (with $B_{tran,mn}$ of the default condition ~0.55), and little nutrient stress. Niwot Ridge indicated substantial amounts of drought ($B_{tran,mn}$ ~0.5), and N limitation, spending 27\% of carbon on nutrient acquisition. Harvard Forest site had a moderate drought stress ($B_{tran,mn}$~90\%), and moderate N limitation (~15\% of assimilation spent on uptake). The Metolius site had some drought stress ($B_{tran,mn}$ ~86\%) and very little nutrient stress.

These differences in the initial conditions at the sites have implications for the parameteric responses, as documented for each parameter in turn in the section below.

@@

For GPP, the impact of SLA is always positive, given increasing LAI, and despite higher LAI causing lower N per unit area and leaves, reducing photosynthetic capacity. The impact of SLA on NPP can however be negative (e.g. Metolius), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance. At Niwot Ridge, an optimal SLA is observed with respect to NPP, again reflecting the assimilation/respiration cost trade-off.

Impact of \emph{froot\_leaf}\\
Root:leaf ratio (\emph{froot\_leaf}) has 1) a direct effect on LAI, via changes in allocation fractions to different tissues, 2) an indirect effect on NPP, given the modified maintenance costs of the resultant fine root biomass pools, 3) affects the uptake of water via root biomass and 4) of nutrients via root-mediated n uptake costs.  The first two impacts increase NPP with \emph{froot\_leaf} and the latter two decrease it.  Both types of effect are seen here, with the Niwot Ridge and Metolius sites (which have the lowest productivity and LAI, and thus presumably very low root biomass in the first place) suffering very large reductions in GPP, NPP, and LAI at low \emph{froot\_leaf} ratios.  Harvard forest and Caxiuana both show increases in all variables.  In Figure \ref{btran state}, we can see that low \emph{froot\_leaf} is a primary control over drought stress at the low productivity sites.

@@

For NPP, the corresponding decreases in respiration rates caused by high \emph{leafcn}, and reduced expenditure on N uptake (Figure \ref{btran state}) somewhat ameliorated the GPP increases, particularly at Caxiuana, which has high ambient temperatures and low carbon use efficiency as a result. At Niwot, an optimum leaf CN ratio was apparent. For NPP this was around the default value, but for LAI it was 0.5s.d. lower, reflecting the complex interactions between the expenditure on N uptake and acquisition of N for tissue growth.

@@

Impact of \emph{medlyn\_slope}\\
The change in the stomatal efficiency parameter of the Medlyn 2011 stomatal conductance model most directly impacts GPP, with cascading impacts on all downstream properties (including on leaf temperature and water stress). At the Caxiuana site (Figure \ref{CAX state}), the very lowest values caused a complete death of the ecosystem. In figure 
\ref{btran state} we can see that low values of the Medlyn stomatal slope ameliorated plant drought stress (by reducing stomatally-mediated water losses). It is worth noting that these same simulations typically had extremely low photosynthesis, particularly at Caxiuana and Metolius.

@@

Impact of \emph{lmr\_intercept}\\
The variance in the reported intercepts in \cite{atkin2016} also have important consequences for ecosystem productivity, in line with the uncertainties in the growth respiration parameterization. At Caxiuana, the warmest site here, the \emph{lmr\_intercept} was one of the most important contributing factors to the NPP, LAI and vegetation carbon states.

@@

Impact of \emph{cn\_flex\_a}\\
\emph{cn\_flex\_a}  is the intercept above which an increase in the cost of N will reduce expenditure on N uptake (see \href{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html#modifications-to-allow-variation-in-c-n-ratios} for details). Thus, low values of \emph{cn\_flex\_a}  results in reductions in expenditure on N uptake when N is too expensive, and therefore on lower tissue N content, and vice-versa.   For the range investigated here, this parameter had limited impact on the system state.

@@

Impact of \emph{cn\_flex\_c}\\
 \emph{cn\_flex\_c} affects the degree to which plants attempt to maintain tissue CN values near to their target when environmental costs are varying. For the state impacts here, the role of this parameter was limited, but for the step-change fertilizations it was much more important, so we will discuss it's role further in the section below.

@@

For the default parameterization, the impacts of CO$_{2}$ fertilization on  GPP (1.44; CAX, 1.44 NWR, 1.24 HVF, 1.45; MET) were fairly constant across sites, but for NPP (1.66; CAX, 1.28; NWR, 1.22; HVF, 1.62; MET) the sites with higher responses had higher N limitation in the control cases.  For LAI (1.47; CAX, 1.24 NWR, 1.25; NVF, 1.27 MET) growth at the Metolius site. Leaf N concentrations also dropped the furthest at this site (0.92; CAX, 0.93; NWR, 0.93 HVF, 0.90 MET) suggesting that expenditure on N uptake might be limited by environmental costs at this site (thus leaf CN deviates further from the target).

For Nitrogen responses, impacts on  GPP were all slightly positive, (1.02; CAX, 1.04; NWR, 1.12; HVF, 1.13 MET) (albeit lower for the drought-stressed Caxiuana and Niwot sites).  For NPP (0.96; CAX, 0.98; NWR, 1.25; HVF, 1.06 MET) increasing $N_{leaf}$ (1.04; CAX, 1.01; NWR 1.02 NVF; 1.01 MET) caused negative impacts as some sites, but impacts on LAI were all positive, reflecting lower costs of N uptake and a greater capacity to acquire N to build tissues and retain carbon for growth (1.06; CAX, 2.09 NWR, 1.21; HVF, 1.12 MET)

\subsection{Impact of parameters}
Across the four sites, parametric impacts were very varied, when compared to the relatively consistant impact of parameters on the system state. A majority of parameters show no consistent relationship with CO$_{2}$ or N deposition responses, and the range of their impacts varied over an order of magnitude between sites. Thus, the parameter space of the model is complex with regard to these changing initial conditions and fertilization experiments. Again, we go through the parameters one at a time to highlight the primary features of the model response and how they relate to expectations based on model structure.

@@

Impact of \emph{froot\_leaf}\\
Despite it’s  impact on the system state,\emph{froot\_leaf} had very little impact on fertilization responses, except in the case of very low values at Metolius, where it increased the CO2 responsiveness and at Harvard, where low values decrease both responses. At Harvard the low root allocation runs had higher initial LAI, and thus the responsiveness of the system to fertilization is lower since the feedback from LAI to GPP is less pronounced.

Impact of \emph{stem\_leaf}\\}
Low values of \emph{stem\_leaf} ratio caused lower responses to fertilization for GPP and LAI at Caxiuana and Harvard Forest (Figures \ref{GPP CO2 and N respones 2001} and\ref{LAI CO2 and N respones 2001}). Initial LAI was larger in these cases, explaining the decease in fertilization response as above.

@@

 At Caxiuana, Niwot and Harvard forest, high  \emph{N\_costs} resulted in stronger LAI responses to fertilization, since these cases had higher N limitation in the first instance (Figure \ref{btran state}).  There was also a slight increase in the CO$_{2}$ fertilization response at high \emph{N\_costs}, perhaps as this high N limitation was alleviated by extra assimilated carbon.

 For very low \emph{N\_costs}, at Caxiuana and Niwot, N responses increased, counter-intuitively.  In these cases, plants appear to make leaves with higher N content (Figure \ref{Leaf N CO2 and N respones 2001}), as per the algorithm which allows a higher $N_{area}$ to be achieved at low N cost. However, the total amount of N in the system (the balance of N inputs and outputs) remains the same with low N costs, and so these cases do not have higher leaf area. When more total N is added, the leaf area can expand, and thus the N concentration declines.

Impact of \emph{frac\_fixers}\\
While nitrogen uptake costs were important to the fertilization responses, they do not affect the total amount of N in the system, and therefore appeared to play a secondary role to the amount of N fixation available, in many cases.  The fraction of N fixers in the system has a particularly dominant role over the model response to step changes in N and to CO$_{2}$. In line with expectations, for GPP and LAI, all sites displayed a 'trade-off', whereby high N fixer systems responded strongly to CO$_{2}$ but not N, and low fixer systems responded strongly to N but not CO$_{2}$.

For NPP, this result is muted for the Caxiuana and Metolius sites, and indeed at Caxiuana the low N fixer cases had a very negative (-8\%) NPP response (despite a positive 7.6\% GPP response), on account of the model increasing the N content of the leaves by up to 10\% in the lowest N cost case (Figure \ref{CAX ndep}). Nonetheless, given the lower N uptake costs in the fertilized N experiment, leaf area was substantially increased (up to 14\%) in these cases too.

At Niwot Ridge and Metolius, the changing fixation capacity had a very large impact on carbon fertilization, but not on nitrogen response, particularly for GPP. At Metolius, this impact persisted, but at Niwot the fixation capacity did have an impact on LAI. Interestingly, many variables increased NPP at this site, but only \emph{fracfixers} increased LAI, suggesting a limitation in the total available N suppressed the impact of all the other case where NPP was increased, and also that water limitation prevented any greater assimilation of carbon.

Impact of \emph{leafcn}\\
The impact of target \emph{leafcn} on fertilization responses was complex (recalling that this was a significant control over the system state at all sites). t Caxiuana, high values \emph{leafcn} (low leaf  $N_{area}$) had much stronger responses to N fertilization, indicating the dominant impact of the carboxylation capacity at this tropical site.  At Niwot Ridge, low values of \emph{leafcn} (low leaf  $N_{area}$) had much stronger responses to CO$_{2}$ fertilization (the strongest responses at this site for GPP), potentially since the low stomatal conductance at this water-limited site could benefit most from additional CO$_{2}$ where carboxylation capacity was high.

Impact of \emph{grperc}\\
\emph{grperc}, or the degree to which growth is 'taxed' is constant and not responsive to any of the physiological processes upstream that depend on fertilization. Thus, its effects are all likely due to indirect impacts on the initial state. Two main impacts were observed. At Niwot Ridge low values of \emph{grperc} caused higher responses to CO$_{2}$ fertilization for GPP and NPP, but not LAI. These values corresponded to high initial LAI (on account of larger amounts of C being available for N uptake) and therefore allowed greater leverage of the increasing CO$_{2}$.  At Harvard forest, high \emph{grperc} had caused low levels of LAI, resulting in increased capacity for CO$_{2}$ fertilization at that site.

Impact of \emph{medlyn\_slope}\\
At Niwot Ridge, low values of the Medlyn Slope parameter (with high water use efficiency) were able to respond very strongly to increasing Nitrogen. These cases also had higher GPP in the initial state, and their higher WUE enabled them to take advantage of the increasing N availability.   These cases also ended up with higher vegetation carbon.

Surpriingly, given its critical role in the responsiveness of stomata to CO$_{2}$, we did not find a widespread impact of the Medlyn parameter on the fertilization reponse, potentially suggesting that the impact on the state condition had already accounted for the impact of this parameter.

@@

Impact of \emph{cn\_flex\_a}, \emph{cn\_flex\_b} and \emph{cn\_flex\_c}\\
As with it's impact on the state, there was limited impact of \emph{cn\_flex\_a} on the fertilization responses, indicating that this parametr is

Impact of \emph{cn\_flex\_b}
Low values of \emph{cn\_flex\_b}, where the leafCN concentration can increase in response to fertilization, showed higher responses to n fertilization, particularly at Harvard Forest. Small values of  \emph{cn\_flex\_b} should  decrease the amount of carbon spent on N uptake when NCosts are high, thus, a relaxation of the cost of N will increase the amount spent on uptake, and subsequently $N_{leaf}$ and all fluxes downstream.  Thus, these changes result from the relaxation of an over-suppression of leaf N content.

>>>>

\begin{document}
%tentative author list...
\authors{Rosie Fisher\affil{1},
William Wieder\affil{1},
Benjamin Sanderson\affil{2},
Charles Koven\affil{3},
Keith Oleson\affil{1},
Chonggang Xu\affil{4},
Joshua Fisher\affil{5},
Mingjie Shi\affil{5},
Anthony Walker\affil{6},
Gordon Bonan\affil{1},
David M. Lawrence\affil{1}
}

@@

For this experiment we utilized six different experimental sites. One in tropical forests (Caxiuan\~a), one in alpine forest (Niwot Ridge), one in temperate forest (Harvard Forest) and one in a seasonally dry evergreen forest (Metolius Forest).

@@

\subsection{Perturbed Physics Ensemble (PPE)}
The CLM5 has a large array of parameters that might have direct and indirect impacts on biogeochemical cycling in the CLM5. For this manuscript, we focus on parameters included in recent developments in the vegetation model, and how they might impact upon the responses to CO$_{2}$ and N fertilization. Therefore, 
we focus on parameters of particular https://www.overleaf.com/project/5b05c31494d9960223796bc3relevance to the C and N cycles and on those which are newly introduced by the modified N cycle. The choice of parameters used here is informed primarily by the model structure, as well as many iterative and informal investigations of the CLM5 model during the development process.

@@

\subsubsection{Parameters Selected for PPE}

@@

\subsubsection{Parameters not included in PPE}

We conducted a set of tests on the parameters of the denitrification model, to assess the impacts of differing rates of N loss on the system. However, recent analysis has shown that the rates of overall denitrification in CLM5 are very low, overall (compared to plant uptake and immobilization \textbf{I've forgotten why this is? Can someone remind me}) and hence these parameters had very little impact. We did not conduct tests on leaf lifespan, since for evergreen trees in CLM5, leaf lifespan and specific leaf area have very similar impacts on overall achievable leaf area, and for deciduous trees, leaf lifespan is primarily determined by phenological triggers. We did not investigate any parameters relating to the crop model (Lombardozzi et al. in prep) nor to wood harvest, fire, dust, or soil carbon decomposition, (the latter is largely unaltered from CLM4.5 as per \cite{koven2013}.

@@

\emph{Specific Leaf Area} and \emph{leafcn} are defined in the TRY plant trait database(\url{www.try-db.org} and we take their values from the analysis by (\cite{kattge2011}) of their mean and distribution. For both parameters, values are log-normally distributed. We here take the PFT average log-normalised standard deviation range from TRY and apply it across all PFTs (thus the lower bounds are closer to the mean than the higher bounds, Table \ref{table_ranges}). The range of \emph{fracfixers} and \emph{frac\_ecto} were set to vary across the whole logical range from 0 to 1 (0 to 100\% of ecosystem productivity being available for fixation or ectomycorrhizal uptake). \emph{grperc} ranges from its logical lower bound of zero up to the range of the previously used CLM value (0.3). \emph{lmr\_intercept} and \emph{medlyn\_slope} are both set such that the range of values tested represents the range across all the PFTs reported in \cite{atkin2015} and \cite{dekauwe2015}. Reasonable ranges of the \emph{cn\_flex} parameters are unknown, since these physiological controls are poorly understood. Thus, we explore an order of magnitude in each direction from the default, to obtain an understanding of the general sensitivity to this parameterization (thus, these results should not be interpreted as covering a constrained range of variation). The same issues apply to $N_{costs}$ parameters.

@@

\subsubsection{Low leaf area investigations}
In general, we found that leaf area indices predicted by the CLM5 default parameter set were often low  compared to published estimates for these sites. (for Caxiuan\~a, 3.0 v.s. 5.3 \cite{fisher2007}, for Niwot Ridge, 1.7 vs. 4.2 \cite{bowling2009}, for Harvard Forest 2.6 vs. 3.5 \cite{williams1996} and for Metolius, 0.5 (modeled) v.s. 1.5 (\cite{spadavecchia2011}). This is a somewhat surprising result, given the generally good agreement of globally-driven model output with LAI reported by \cite{lawrence2018}, and may plausibly reflect differences between the use of reanalysis products used to drive the model and site-specific meteorological data.

To interrogate this further, we ran a separate set of simulations with higher allocation to leaves (generated from lower stem allocation parameters). These simulations indicated that, with the resulting higher, and more consistent with observed, LAI values, the model was less sensitive to both CO$_{2}$ and N deposition, given the amplifying effect of increasing leaf area increasing fluxes at unsaturated LAI levels.   For example, at Harvard Forest, the high leaf allocation  simulation (LAI of 4.6, vs 2.6 for the default parameters) had a lower CO$_{2}$ response (1.43, 1.66, 1.47 for GPP, NPP and LAI in the control vs. 1.18, 1.17 and 1.22 for the high LAI case) and a lower N response as well (1.1, 1.26, 1.21 for the control GPP, NPP, LAI, and 1.05, 1.20 and 1.15 for the high LAI case). Similar results were found at all other sites.  Thus, by virtue of their lower LAI, the overall fertilization rates at these sites may be substantially larger than those in the model driven with reanalysis meteorology (analysis tbd).

\subsubsection{Water and N limitation across sites}
Figure \ref{btran state} illustrates the relative water and nitrogen limitation in the control (unfertilized) parameter ensemble. Water limitation in CLM5 can be expressed via the $B_{tran,mn}$ parameter, which is a model property calculated as a function of prognostic leaf water potential, which impacts gas exchange processes in a similar manner to the $B_{tran}$ parameter in CLM4.5 (in that it is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$). Because leaf water potential changes diurnally, we now use the minimum term to calculate relative water stress between sites. For N limitation, the fraction of total NPP which is used to pay for nitrogen uptake, is a direct indicator of how N stress affects the carbon budget.

The Caxiuan\~a site experienced substantial drought stress in these simulations (with $B_{tran,mn}$ of the default condition ~0.55), and little nutrient stress. Niwot Ridge indicated substantial amounts of drought ($B_{tran,mn}$ ~0.5), and N limitation, spending 27\% of carbon on nutrient acquisition. Harvard Forest site had a moderate drought stress ($B_{tran,mn}$~90\%), and moderate N limitation (~15\% of assimilation spent on uptake). The Metolius site had some drought stress ($B_{tran,mn}$ ~86\%) and very little nutrient stress.

These differences in the initial conditions at the sites have implications for the parametric responses, as documented for each parameter in turn in the section below.

@@

For GPP, the impact of SLA is always positive, given increasing LAI, and despite higher LAI causing lower N per unit area of leaves which reduces photosynthetic capacity. The impact of SLA on NPP can however be negative (e.g., Metolius), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance. At Niwot Ridge, an optimal SLA is observed with respect to NPP, again reflecting the assimilation/respiration cost trade-off.

Impact of \emph{froot\_leaf}\\
Root:leaf ratio (\emph{froot\_leaf}) has 1) a direct effect on LAI, via changes in allocation fractions to different tissues, 2) an indirect effect on NPP, given the modified maintenance costs of the resultant fine root biomass pools, 3) an effect on the uptake of water via root biomass and 4) an effect on nutrient limitation??? through root-mediated n uptake costs.  The first two impacts increase NPP with \emph{froot\_leaf} and the latter two decrease it.  Both types of effect are seen here, with the Niwot Ridge and Metolius sites (which have the lowest productivity and LAI, and thus presumably very low root biomass in the first place) suffering very large reductions in GPP, NPP, and LAI at low \emph{froot\_leaf} ratios.  Harvard forest and Caxiuan\~a both show increases in all variables.  In Figure \ref{btran state}, we can see that low \emph{froot\_leaf} is a primary control over drought stress at the low productivity sites.

@@

For NPP, the corresponding decreases in respiration rates caused by high \emph{leafcn}, and reduced expenditure on N uptake (Figure \ref{btran state}) somewhat ameliorated the GPP increases, particularly at Caxiuan\~a, which has high ambient temperatures and low carbon use efficiency as a result. At Niwot, an optimum leaf CN ratio was apparent. For NPP this was around the default value, but for LAI it was 0.5s.d. lower, reflecting the complex interactions between the expenditure on N uptake and acquisition of N for tissue growth.

@@

Impact of \emph{medlyn\_slope}\\
The change in the stomatal efficiency parameter of the Medlyn 2011 stomatal conductance model most directly impacts GPP, with cascading impacts on all downstream properties (including on leaf temperature and water stress). At the Caxiuan\~a site (Figure \ref{CAX state}), the very lowest values caused a complete death of the ecosystem. In figure 
\ref{btran state} we can see that low values of the Medlyn stomatal slope ameliorated plant drought stress (by reducing stomatally-mediated water losses). It is worth noting that these same simulations typically had extremely low photosynthesis, particularly at Caxiuan\~a and Metolius.

@@

Impact of \emph{lmr\_intercept}\\
The variance in the reported intercepts in \cite{atkin2016} also have important consequences for ecosystem productivity, in line with the uncertainties in the growth respiration parameterization. At Caxiuan\~a, the warmest site here, the \emph{lmr\_intercept} was one of the most important contributing factors to the NPP, LAI and vegetation carbon states.

@@

Impact of \emph{cn\_flex\_a}\\
\emph{cn\_flex\_a}  is the intercept above which an increase in the cost of N will reduce expenditure on N uptake (see \href{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html#modifications-to-allow-variation-in-c-n-ratios} for details). Thus, low values of \emph{cn\_flex\_a}  result in reductions in expenditure on N uptake when N is too expensive, and therefore in lower tissue N content, and vice-versa.   For the range investigated here, this parameter had limited impact on the system state.

@@

Impact of \emph{cn\_flex\_c}\\
 \emph{cn\_flex\_c} affects the degree to which plants attempt to maintain tissue C:N values near to their target when environmental costs are varying. For the state impacts here, the role of this parameter was limited, but for the step-change fertilizations it was much more important, so we will discuss it's role further in the section below.

@@

For the default parameterization, the impacts of CO$_{2}$ fertilization on  GPP (1.44; CAX, 1.44 NWR, 1.24 HVF, 1.45; MET) were fairly constant across sites, but for NPP (1.66; CAX, 1.28; NWR, 1.22; HVF, 1.62; MET) the sites with higher responses had higher N limitation in the control cases.  For LAI (1.47; CAX, 1.24 NWR, 1.25; NVF, 1.27 MET) growth at the Metolius site. Leaf N concentrations also dropped the furthest at this site (0.92; CAX, 0.93; NWR, 0.93 HVF, 0.90 MET) suggesting that expenditure on N uptake might be limited by environmental costs at this site (thus leaf C:N deviates further from the target).

For Nitrogen responses, impacts on  GPP were all slightly positive, (1.02; CAX, 1.04; NWR, 1.12; HVF, 1.13 MET) (albeit lower for the drought-stressed Caxiuan\~a and Niwot sites).  For NPP (0.96; CAX, 0.98; NWR, 1.25; HVF, 1.06 MET), increasing $N_{leaf}$ (1.04; CAX, 1.01; NWR 1.02 NVF; 1.01 MET) caused negative impacts on NPP at some sites, even though impacts on LAI were all positive, reflecting lower costs of N uptake and a greater capacity to acquire N to build tissues and retain carbon for growth (1.06; CAX, 2.09 NWR, 1.21; HVF, 1.12 MET)

\subsection{Impact of parameters}
Across the four sites, parametric impacts were very varied, when compared to the relatively consistent impact of parameters on the system state. A majority of parameters show no consistent relationship with CO$_{2}$ or N deposition responses, and the range of their impacts varied over an order of magnitude between sites. Thus, the parameter space of the model is complex with regard to these changing initial conditions and fertilization experiments. Again, we go through the parameters one at a time to highlight the primary features of the model response and how they relate to expectations based on model structure.

@@

Impact of \emph{froot\_leaf}\\
Despite its strong impact on the system state,\emph{froot\_leaf} had very little impact on fertilization responses, except in the case of very low values at Metolius, where it increased the CO2 responsiveness and at Harvard, where low values decrease both responses. At Harvard the low root allocation runs had higher initial LAI, and thus the responsiveness of the system to fertilization is lower since the feedback from LAI to GPP is less pronounced.

Impact of \emph{stem\_leaf}\\}
Low values of \emph{stem\_leaf} ratio caused lower responses to fertilization for GPP and LAI at Caxiuan\~a and Harvard Forest (Figures \ref{GPP CO2 and N respones 2001} and\ref{LAI CO2 and N respones 2001}). Initial LAI was larger in these cases, explaining the decease in fertilization response as above.

@@

 At Caxiuan\~a, Niwot and Harvard forest, high  \emph{N\_costs} resulted in stronger LAI responses to fertilization, since these cases had higher N limitation in the first instance (Figure \ref{btran state}).  There was also a slight increase in the CO$_{2}$ fertilization response at high \emph{N\_costs}, perhaps as this high N limitation was alleviated by extra assimilated carbon.

 For very low \emph{N\_costs}, at Caxiuan\~a and Niwot, N responses increased, counter-intuitively.  In these cases, plants appeared to have leaves with higher N content in the initial state (Figure \ref{NR1 state}), as per the algorithm which encourages a higher $N_{area}$ to be at lower N costs.  This appears to restrict the total LAI in the initial state, given that, the total amount of N in the system (the balance of N inputs and outputs) is not changed by \emph{N\_costs}, and so these cases do not have higher leaf area. When more total N is added, the leaf area can expand, and thus the N concentration declines.

Impact of \emph{frac\_fixers}\\
While nitrogen uptake costs are important to the fertilization responses, they do not affect the total amount of N in the system, and therefore appeared to play a secondary role to the amount of N fixation available.

The fraction of N fixers in the system has a particularly dominant role over the model response to step changes in N and to CO$_{2}$. In line with expectations, for GPP and LAI, all sites displayed a 'trade-off', whereby high N fixer systems responded strongly to CO$_{2}$ but not N, and low fixer systems responded strongly to N but not CO$_{2}$ (Figure \ref{GPP CO2 and N respones 2001}, \ref{LAI CO2 and N respones 2001}).

For NPP, N deposition responses were muted at Metolius, and reversed in sign at Caxiuan\~a, where the low N fixer cases had a very negative (-8\%) NPP response (despite a +7.6\% GPP response), on account of the model increasing the N content of the leaves by up to 10\% in the lowest N cost case (Figure \ref{CAX ndep}). Nonetheless, given the lower N uptake costs in the fertilized N experiment, leaf area was substantially increased (up to 14\%) in these cases too.

At Niwot Ridge the changing fixation capacity had no impact on the response of GPP to N fixation, despite its influence on LAI, suggesting that water limitation was a greater constraint on total GPP than leaf area.

At Metolius, again the changing fixation capacity had no impact on the response of GPP to N deposition, but in this case, it had no impact on LAI either. Only SLA had a large impact on N fertilization at this site, suggesting that only the higher LAI runs were able to respond to fertilization.

Impact of \emph{leafcn}\\
The impact of target \emph{leafcn} on fertilization responses was complex (recalling that this was a significant control over the system state at all sites).

At Caxiuan\~a, high values of \emph{leafcn} (low leaf $N_{area}$) had much stronger responses to N fertilization, indicating the dominant impact of carboxylation rate limitation at this tropical site.

At Niwot Ridge, low values of \emph{leafcn} (high leaf $N_{area}$) had much stronger responses to CO$_{2}$ fertilization (the strongest response at this site for GPP). These cases suffered from high water limitation, given their large fluxes (Figure \ref{btran state}) and thus the CO$_{2}$ fertilization would alleviate this problem.

Impact of \emph{grperc}\\
\emph{grperc}, or the degree to which growth is 'taxed' is constant and not responsive to any of the physiological processes upstream that depend on fertilization, thus, at most sites, impacts are minimal. At Niwot Ridge low values of \emph{grperc} caused higher responses to CO$_{2}$ fertilization for GPP and NPP, and LAI.  These values corresponded to high initial LAI  and therefore allowed greater leverage of the increasing CO$_{2}$.

Impact of \emph{medlyn\_slope}\\
At Niwot Ridge, low values of the Medlyn Slope parameter (with high water use efficiency) were able to respond very strongly to increasing Nitrogen (Figure \ref{NR1 ndep}, given their lower initial water stress (Figure \ref{btran state}).

Surprisingly however, given its critical role in the responsiveness of stomata to CO$_{2}$, we did not find a widespread impact of the Medlyn parameter on the fertilization response, potentially suggesting that the impact on the state condition had already accounted for the impact of this parameter.

@@

Impact of \emph{cn\_flex\_a}\\
As with it's impact on the state, there was limited impact of \emph{cn\_flex\_a} on the fertilization responses.

Impact of \emph{cn\_flex\_b}
At Harvard Forest, low values of  \emph{cn\_flex\_b} increased responses to N deposition and decreased response to CO$_[2}$. Leaf N concentration followed suit, increasing substantially in the N deposition experiment.

Low values of \emph{cn\_flex\_b} decrease the amount of carbon spent on N uptake when \emph{N\_costs} are high, therefore potentially suppressing leaf N content. Adding N overcomes this suppression, leading to greater fertilization responses. (n.b. I'm not sure about this, need to come back to it again).

Impact of \emph{cn\_flex\_c}\\
The impact of \emph{cn\_flex\_c} (the parameter controlling the degree to which plant C:N ratios are constrained to their target) appeared to have its strongest impact on the fertilization responses of vegetation carbon (Figure \ref{VEGC CO2 and N respones 2001})  at Harvard Forest and Niwot Ridge (and Caxiuan\~a, to a smaller extent). These impacts do not mirror those on leaf area index or NPP, and overall the response to this parameter appears complex.

At Metolius and Niwot, this parameter has a dominant control over the impacts of CO$_{2}$ fertilization on $N_{leaf}$ (Figure \ref{Leaf N CO2 and N respones 2001}).  The impacts on vegetation carbon are likely due to the larger decreases in tissue N content that are permissable when values of this parameter are high (e.g. Figure \ref{NR1 celev}, \ref{MET celev}).  Low  $N_{leaf}$ values permit higher leaf area and vegetation carbon, but reduce assimilation rates, as illustrated at Harvard Forest and Niwot Ridge (Figure \ref{HRV celev}, \ref{CAX celev}).

\section{Discussion}
The overall responses of land surface models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP model inter-comparison projects and associated activities  (\cite{meehl2014}).  Throughout the CMIP activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space. At the time of writing, we do not yet know the climate feedback strength of the default version of CLM5 compared to it's predecessors, but it is clear from these investigations that many model parameters can and will have had a large impact on these outcomes.
